Demes: a specification for interchanging complex demography

### 2021-06-24 paper.tex
start at pop gen model section

In Demes, demographic models consist of one or more interacting
% JK: not sure at all about this, but let's see how it goes for now.
populations (or ``demes"; to avoid confusion with the name of the
specification itself we will use the term ``population'' in this
discussion, in the understanding that the terms are interchangeable
for our purposes). Populations are defined as a collection of
exchangable individuals that exist for some period of time,
following a well-defined set of rules regarding population
sizes, mating systems, etc. Ancestor/descendent relationships
between populations can be defined, as well as
possible continuous or instantaneous migration between coexisting
populations. The history of each population is defined as a
series of ``epochs'' during which the population's parameters are
fixed.

@@

>>>>

@@

\subsection*{Population genetics model}

In Demes, demographic models consist of one or more interacting populations (or
``demes"; to avoid confusion with the name of the specification itself we will
use the term ``population'' in this discussion, in the understanding that the
terms are interchangeable for our purposes). Populations are defined as a
collection of exchangable individuals that exist a specified period of time,
following a well-defined set of rules regarding population sizes, mating
systems, etc. Ancestor/descendent relationships between populations can be
defined, as well as continuous or instantaneous migration between coexisting
populations.

Population and event times are written as units in the past, so that time zero
corresponds to the final generation or ``now'', and event times in the past are
values greater than zero with larger values for events that occur in the more
distant past. While a natural specification for time units is in generations,
other time units are allowed, such as years. In the case the the time units are
not given in generations, the generation time must also be specified so that
times can be converted into generations. In general, as time flows from the
past to the present, populations, epochs, and migration events should be
specified in their order of appearance, so that their times are in descending
order.

Models include one or more populations that have sizes greater than zero for
the duration of their existences, and population sizes are given as the number
of individuals. A population must exist for some duration of time greater than
zero, so that its start time must be larger than its end time. It may have one
or more ancestors, which are other populations that exist at the population's
start time. If one ancestor is specified, the first generation is constructed
by randomly sampling parents from the ancestral population to contribute to
offspring in the newly generated population. If more than one ancestor is
specified, the proportions of ancestry from each contributing population must
be provided, and those proportions must sum to one. In this case, parents are
chosen randomly from each ancestral population with probability given by those
proportions. If no ancestors are specified, the population is assumed to have
start time equal to infinity, as individuals are not assumed to spontaneously
spring into existence.

Within populations, population sizes and mating systems are specified within
epochs. Epochs are non-overlapping intervals that cover the entire existence
interval of the population. Each epoch specifies the population size over that
interval, which can be a constant value or function defined by start and end
sizes that must remain positive.  If an epoch has a start time of infinity, the
population size for that epoch must be constant. Epochs can also specify
parameters for nonrandom mating, such as selfing or cloning rates, which give
the probability that offspring are generated from one generation to the next by
self-fertilization or cloning of an indivual \textbf{need to precisely specify
once this is determined}.

Finally, individuals in a population may have parents from a different
population through migrations. These can be defined as continuous migration
rates over intervals that populations overlap in time or through instantaneous
(or pulse) migration events at a given time. Continuous migration rates are
defined as the probability that parents in the ``destination`` population are
chosen from the ``source`` population.  Migration rates are thus per-generation
and must be less than one. Furthermore, if more than one source population have
continuous migration into the same destination population, the sum of those
migration rates must also be less than one, as rates define probabilities. The
probability that parents come from the same population is just one minus the
some of incoming migration rates. On the other hand, pulse migration events
specify the instantaeous replacement of a given fraction of individuals in a
destination population by individuals with parents from a source population.
The fraction must be between zero and one, and if more than one source
population has a pulse migration event to the same destination population and
the same time, those replacement events are applied sequentially in the order
that they are specified in the model.
